Here is the file with the examples and solutions formatted:

```{r setup, echo=FALSE}
source("intro-statistics-quarto.R")
```
# Backwards Problems {#backwards-problems}

In this section we want to go in the opposite direction from previous chapters. That is we want to find an **_x-value that will give us a certain left tail or right tail area_**. These are called **backwards** problems since they start with the area and then find the x-value. This is backwards from what we have done before.

So in previous chapters we had this:

_**Find Area From an X-value**_

1. Start with an x-value 
1. Compute its z-value
1. Find the area using the z table (or use NORMSDIST)

Now we want to go in the other direction:  

_**Find X-value From an Area**_

1. Start with an area 
1. Find the z-value that gives that area
1. Compute the x-value from the z-value 

Before we do these full backwards problems from area to x-value, first lets solve an easier problem that involves finding the z-value that goes with a particular area.  

Then once we can do that, we will see how to find the x-value from the z-value for backwards problems.  
    
Okay so we are given an area (left or right tail) and we want to find the z-value.

## Find Z Value for Given Left Tail Area 

:::{#exm-z-value-for-left-tail-of-25%}
## z-value for left tail of $25\\%$
```{r}
knit_example(list(area=0.25), 'BackwardsStdNormalLeftQuestion.Rmd', 'BackwardsStdNormalLeftSolution.Rmd')  
```
:::

## Find Z Value for Given Right Tail Area

:::{#exm-z-value-for-right-tail-of-18%}  
## z-value for right tail of $18\\%$
```{r}
knit_example(list(area=0.18), 'BackwardsStdNormalRightQuestion.Rmd', 'BackwardsStdNormalRightSolution.Rmd')
```
:::

## Solve Backwards for X-value

Once you have the z-value that you need using one of the above you can solve backward for the x-value that you need. It works like this: 

_**Find X-value by Solving Z-score Formula Backwards**_

1. Find the z-value you need using technique above
1. Plug the z-value into the z-score formula  
1. Plug in $\mu$ and $\sigma$ (these are known)
1. Solve for the x-value

So basically you just plug your z-value, $\mu$ and $\sigma$ into this and solve for x:

\begin{equation} 
z =\frac{x-\mu}{\sigma} 
\end{equation}

:::{#exm-Find-$x$-for-bottom-25%-of-data}  
## Find $x$ for bottom $25\\%$ of data
```{r}
knit_example(list(area=0.25, mu=18, sigma=3), 'SolveBackwardsLeftTailQuestion.Rmd', 'SolveBackwardsLeftTailSolution.Rmd')
```
:::

:::{#exm-Find-$x$-for-top-15%-of-data}
## Find $x$ for top $15\\%$ of data
```{r}
knit_example(list(area=0.15, mu=23, sigma=4), 'SolveBackwardsRightTailQuestion.Rmd', 'SolveBackwardsRightTailSolution.Rmd')
```
:::

## Applications of Backwards Problems

:::{#exm-Bottom-10%-of-Demand-Values}
## Bottom $10\\%$ of Demand Values
```{r} 
knit_example(list(area=0.10, mu=257, sigma=45), 'BottomPercentDemandQuestion.Rmd', 'BottomPercentDemandSolution.Rmd')
```
:::

Here is another application of backwards problems:

The **_service level_** in retail corresponds to the chance that a retailer would experience a lost sale (out of stock situation) for some period of interest.

- A service level of 90% means that lost sales would happen 10% of the time.   
- A service level of 95% means that lost sales would happen be 5% of the time. 
- A service level of 99% means that lost sales would happen be 1% of the time. 

Different items and categories can have different service levels as well, with some items being "high" service level items while others being "low" service level items.

:::{#exm-Inventory-For-95%-service-level}
## Inventory For $95\\%$ service level
```{r}
knit_example(list(area=0.95, mu=340, sigma=80), 'InventoryLevelQuestion.Rmd', 'InventoryLevelSolution.Rmd')
```  
:::

In this interpretation service level corresponds to a certain left tail area for the normal distribution:  

- A service level of 90% means you want a left tail area of 90%
- A service level of 95% means you want a left tail area of 95%.
- A service level of 99% means you want a left tail area of 99%.

Or you can say it this way as well: 

- A service level of 90% means you have lost sales 10% of the time.   
- A service level of 95% means you have lost sales 5% of the time.
- A service level of 99% means you have lost sales 1% of the time.
