```{r setup, echo=FALSE}
source("intro-statistics-quarto.R")
```
# Backwards Problems {#backwards-problems}

In this section we want to go in the opposite direction from previous chapters. That is we want to find an **_x-value that will give us a certain left tail or right tail area_**. These are called **backwards** problems since they start with the area and then find the x-value. This is backwards from what we have done before.

So in previous chapters we had this:

_**Find Area From an X-value**_

1. Start with an x-value
1. Compute its z-value
1. Find the area using the z table (or use NORMSDIST)

Now we want to go in the other direction:

_**Find X-value From an Area**_

1. Start with an area
1. Find the z-value that gives that area
1. Compute the x-value from the z-value

Before we do these full backwards problems from area to x-value, first lets solve an easier problem that involves finding the z-value that goes with a particular area.

Then once we can do that, we will see how to find the x-value from the z-value for backwards problems.

Okay so we are given an area (left or right tail) and we want to find the z-value.

:::{#exm-z-value-for-left-tail-z-equals-1.23}
## z-value for left tail $z=1.23$
```{r}
knit_example(list(area=0.25),
                  'statplosion-problems/question/BackwardsStdNormalLeftQuestion.Rmd',
                  'statplosion-problems/solution/BackwardsStdNormalLeftSolution.Rmd')
```
:::

:::{#exm-z-value-for-right-tail-of-18-percent}
## z-value for right tail of $18\%$
```{r}
knit_example(list(area=0.18),
                  'statplosion-problems/question/BackwardsStdNormalRightQuestion.Rmd',
                  'statplosion-problems/solution/BackwardsStdNormalRightSolution.Rmd')
```
:::

## Solve Backwards for X-value

Once you have the z-value that you need using one of the above you can solve backward for the x-value that you need. It works like this:

_**Find X-value by Solving Z-score Formula Backwards**_

1. Find the z-value you need using technique above
1. Plug the z-value into the z-score formula
1. Plug in $\mu$ and $\sigma$ (these are known)
1. Solve for the x-value

So basically you just plug your z-value, $\mu$ and $\sigma$ into this and solve for x:

$$
z =\frac{x-\mu}{\sigma}
$$ {#eq-solve-this-for-x}

Here is an example where we want to find the **lower 25%** of the data. That means we want to find the x-value that has a left tail area of 25%.

:::{#exm-Find-x-for-bottom-25-percent-of-data}
## Find $x$ for bottom $25\%$ of data
```{r}
knit_example(list(area=0.25, mu=18, sigma=3),
                  'statplosion-problems/question/SolveBackwardsLeftTailQuestion.Rmd',
                  'statplosion-problems/solution/SolveBackwardsLeftTailSolution.Rmd')
```
:::

Here is an example where we want to find the **upper 15%** of the data. That means we want to find the x-value that has a right tail area of 15%.

:::{#exm-Find-x-for-top-15-percent-of-data}
## Find $x$ for top $15\%$ of data
```{r}
knit_example(list(area=0.15, mu=23, sigma=4),
                  'statplosion-problems/question/SolveBackwardsRightTailQuestion.Rmd',
                  'statplosion-problems/solution/SolveBackwardsRightTailSolution.Rmd')
```
:::

## Applications of Backwards Problems

Here is an example where you want to find out when a demand value would be in the bottom 10% of your demand expectations for an item.

:::{#exm-Bottom-10-percent-of-Demand-Values}
## Bottom $10\%$ of Demand Values
```{r}
knit_example(list(area=0.10, mu=257, sigma=45),
                  'book-problems/question/BottomPercentDemandQuestion.Rmd',
                  'book-problems/solution/BottomPercentDemandSolution.Rmd')
```
:::

Here is another application of backwards problems:

The **_service level_** in retail corresponds to the chance that a retailer would experience a lost sale (out of stock situation) for some period of interest.

- A service level of 90% means that lost sales would happen 10% of the time.
- A service level of 95% means that lost sales would happen be 5% of the time.
- A service level of 99% means that lost sales would happen be 1% of the time.

Different items and categories can have different service levels as well, with some items being "high" service level items while others being "low" service level items.

The next example shows how to set inventory levels for an item to achieve a certain service level.

:::{#exm-Inventory-For-95-percent-service-level}
## Inventory For $95\%$ service level
```{r}
knit_example(list(area=0.95, mu=340, sigma=80),
                  'book-problems/question/InventoryLevelQuestion.Rmd',
                  'book-problems/solution/InventoryLevelSolution.Rmd')
```
:::

In this interpretation service level corresponds to a certain left tail area for the normal distribution:

- A service level of 90% means you want a left tail area of 90%
- A service level of 95% means you want a left tail area of 95%.
- A service level of 99% means you want a left tail area of 99%.

Or you can say it this way as well:

- A service level of 90% means you have lost sales 10% of the time.
- A service level of 95% means you have lost sales 5% of the time.
- A service level of 99% means you have lost sales 1% of the time.


