```{r setup, echo=FALSE}
source("intro-statistics-quarto.R")
library("statplosion")
options(scipen=999)
```

```{r}
barchart<-function(x,
                   names,
                   xlab="x",
                   ylab="y",
                   main="Title") {
    ymax <- max(x)
    g<-barplot(x,
            names.arg=names,
            las=1,
            xlab=xlab,
            ylim=c(0,ymax*1.1),
            ylab=ylab,
            main=main)
    text(g, x, labels=x, pos=3)
}
```


# Barplots {#barplots}

Most people are familar with bar charts. They show amounts for different categories.

The amounts will typically be either counts (frequencies) or percents.

We look at an example of both here.


## Market Share

Suppose the following fashion stores having competing products with the following market shares:

```{r}
sales<-c(200000,100000,600000,300000)
company<-c("Pop Shop","Tommy Go Figure","Never 42","Country Infitters")
df<-data.frame(Company=company, Sales=sales)
kable(df, "markdown")
```
Here is the barplot that goes with this:

```{r, example=T, title="Market Share By Dollars", fig.align="center"}
barchart(sales,company,xlab="",ylab="", main="Market Share By Dollars")
```

Now suppose we want the percentage breakdown of these market shares:

- From `r company[1]`:

$$
\frac{`r sales[1]`}{`r sum(sales)`} =
`r round(sales[1]/sum(sales)*100)`\%
$$

- From `r company[2]`:

$$
\frac{`r sales[2]`}{`r sum(sales)`} =
`r round(sales[2]/sum(sales)*100)`\%
$$

- From `r company[3]`:

$$
\frac{`r sales[3]`}{`r sum(sales)`} =
`r round(sales[3]/sum(sales)*100)`\%
$$

- From `r company[4]`:

$$
\frac{`r sales[4]`}{`r sum(sales)`} =
`r round(sales[4]/sum(sales)*100)`\%
$$

So here is the market share by percentage:

```{r}
s<-sum(sales)
percents <- round(100*sales/s)
percentsdf<-data.frame(Company=company, Percent=percents)
kable(percentsdf, "markdown")
```

Now if we make a barplot with percentages this looks like this:

:::{#exm-market-share-by-percent}
## Market Share By Percent
```{r, fig.align="center"}
barchart(percents,company, xlab="",ylab="percents %", main="Market Share By Percent")
```
$$ \tag*{$\blacksquare$} $$
:::

Of course the graph does not look any different, but it is nice to see the percentage breakdown in the chart easily.

```{r}
knit_exit()
```

## Market Share with Competitor

Now suppose a competitor "Country Infitter" is going to introduce a product that competes with the above.

Suppose it is expected they will generate \$100,000 and will take 3/4 of their revenues from the existing stores.

We assume they will take dollar amounts based on their percents:


- From `r company[1]`:
$$
.25(`r sales[1]`)=`r .25*sales[1]`
$$
- From `r company[2]`:
$$
.25(`r sales[2]`)=`r .25*sales[2]`
$$
- From `r company[3]`:
$$
  .25(`r sales[3]`)=`r .25*sales[3]`
$$

from each of the companies.

```{r}
new<-.75*sales
new<-c(new,100000)
company<-c(company, "Country Infitter")
```

:::{#exm-market-share-by-dollar}
## Market Share By Dollar
```{r, fig.align="center"}
barchart(new,company,xlab="",ylab="dollars $", main="Market Share By Dollars")
```
$$ \tag*{$\blacksquare$} $$
:::

And here is the new market share given by percents

```{r}
s<-sum(new)
percents <- round(100*new/s)
percentsdf<-data.frame(Company=company, Percent=percents)
kable(percentsdf, "markdown")
```

```{r, example=T, title="New Market Share by Percent", fig.align="center"}
barchart(percents,company,xlab="",ylab="percents %", main="Market Share By Percent")
```
