```{r setup, echo=FALSE}
source("intro-statistics-quarto.R")

```
# Confidence Intervals For Means {#confidence-intervals-means}

In this section we look at confidence intervals for a population mean $\mu$. 

This is a technique for estimating the population mean when we do not know it. 

Use this technique when you have a sample mean and and want to use it to estimate what the population mean is. 

This is an example of inferential statistics since we are using information from a sample (the statistic $\bar x$) to infer information about the population (the parameter $\mu$).
    
We start by taking a sample and finding a sample mean $\bar x$. Then we add and subtract a margin of error $E$ to the sample mean to make an interval where we hope to find the population mean. We will show how to calculate the margin of error $E$ below. 

## Steps for Finding Confidence Intervals - Means

The steps involved to do this are as follows:

- Take a sample and calculate the sample mean $\bar x$
- Calculate a margin of error $E$ using the formula below. (This tells us roughly how far off $\mu$ is from $\bar x$)
- Find the interval by adding and subtracting $E$ to $\bar x$. (This interval tells where we hope to find $\mu$)

```{asis example=TRUE, title="A Confidence Interval and Margin of Error", echo=TRUE}

Suppose we take a sample and collect some data. Then we compute a sample mean: 

$$
\bar x = 131 
$$

Then using the formula for a confidence interval suppose we find that 

$$
E = 22 
$$

(We will show how to compute this later on. For now just pretend we did it and got $22$.)

Subtract $E$ from $\bar x$ to get the **lower endpoint** of the confidence interval:

$$
\bar x - E = 131 - 22 = 109 
$$

Add $E$ to $\bar x$ to get the **upper endpoint** of the confidence interval:

$$
\bar x + E = 131 + 22 = 153
$$

So the confidence interval goes from 109 up to 153. Sometimes we write this like this:

$$
(\bar x - E,\ \bar x + E) = (109,153)
$$

Roughly what this says is that we believe the population mean is **between 109 and 153**. 

Exactly how confident and in what sense the above is true we will see below.
```

Next we look at how we come up with $E$, the margin of error.

## Margin of Error E for Confidence Intervals - Means

To find $E$ we have to first decide on a **confidence level.** This is a percentage (something like 90% or 95% or 99%). 

The confidence level describes how confident we are that our confidence intervals will contain the population mean: 

- A confidence level of 95% means that our confidence intervals ($\bar x - E$, $\bar x + E$) will contain the 
population mean $\mu$ about 95% of the time.
- A confidence level of 99% means that our confidence intervals ($\bar x - E$, $\bar x + E$) will contain the 
population mean $\mu$ about 95% of the time.
- and so on

Another way to put this:

For 95% confidence, this means that if we repeatedly take samples and find a sample mean $\bar x$, we can expect to find the population mean inside our confidence intervals roughly 95% of the time.    
    
This notion of confidence level makes sense only in the context of repeatedly taking confidence intervals for a particular situation.  
    
Of the different confidence levels, by far **the most commonly used confidence interval is one with confidence level 95%.** 
    
Once we decide on the confidence level we want to use, we will use the following formula to calculate the margin of error E:

$$
E=(t^\star)\frac{s}{\sqrt{n}} 
$$

The $s$ is the sample standard deviation, the $n$ is the sample size and the $t^\star$ comes from the chosen confidence level. We show where to get it in the next section.

## Critical Value $t^\star$

The formula for E has something called $t^\star$ in it. This is called the **critical value**. It comes directly from the confidence level. Once you know the confidence level you are using you can use the following formula for $t^\star$ that goes with that confidence level $c$. Here is what you should use in a spreadsheet: 

**=TINV(1.0-c, n-1)**

where $c$ is the confidence level as a decimal (so for example 95% would be .95) and $n$ is the sample size. 

## Confidence Interval Calculations - Means 

:::{#exm-a-99-percent-confidence-interval-for-means}
## A $99\%$ Confidence Interval for Means
```{r}
knit_example(list(xbar=12.4,s=5.1, n=38,confidence=0.99),
	     'statplosion-problems/question/ConfidenceIntervalMeansQuestion.Rmd', 
	     'statplosion-problems/solution/ConfidenceIntervalMeansSolution.Rmd') 
```
:::

