```{r setup, echo=FALSE}
source("intro-statistics-quarto.R")
library("statplosion")
```

# Hypothesis Testing - Proportions {#hypothesis-testing-proportions}

Remember we said that typically we do not know things like parameters
for populations. For example we don't always know population proportions
in situations we are interested in.

But suppose **someone claims that a population proportion is some
value** (like 75%).

What kinds of sample proportions would we expect to get from such a
population?

If you remember the Central Limit Theorem it says that in this case the
sampling distribution of sample proportions would look something like
this:

```{r fig.height=2, fig.width=3, fig.align="center"}
p <- 0.75
stdev <- sqrt(p*(1-p)/120)
normal_curve(p,round(stdev,4), show=FALSE, axis_label=bquote(hat(p)))
```
with the middle of it at .75.

So if this were true, if we took sample proportions we would usually
wind up near .75, and more rarely we would get sample proportions that
would be farther away.

We can use this idea to **test the claim that the population proportion
is equal to the claimed value**.

This is called a **hypothesis test** of the (claimed) population
proportion. The results from a hypothesis test will be roughly this:

- If our **sample proportion is too far away** we say it is **evidence against the claim**.
- If our **sample proportion is close to the middle**, we say it is
**no evidence against the claim**.

Basically if the sample proportion is too far away from the (claimed)
population proportion, we believe that claim about the population
proportion may be bogus.

Now remember someone can claim anything, even something outrageous\....
Like "90% of Republicans will vote for Obama" or "FIT consists of 50%
male students".

With Hypothesis Testing we can test the validity of such claims (and
more reasonable claims too).

By looking at a sample, and examining how close the sample proportion is
to the (claimed) population proportion we can figure out whether the
claim makes sense or not.

## Steps for Hypothesis Testing

Note: We use $p_{0}$ as the notation for the proposed population proportion. You should think of this as an actual number like $.32$ or $.75$. It is the **claimed** population proportion.  

### 1) Write down the claimed null hypothesis {-}

$H_{0}:\ p = p_{0}$

This is called null hypothesis (or "H zero") and its interpretation in
words is this:

"**The population proportion is equal to $p_{0}$**"

### 2) Write down the alternative hypothesis {-}

This will be one of the following types:

-   $H_{a}:\ p < p_{0}$ (This one is called a **left tail test** )

This one is interpreted "**the population proportion is less than $p_{0}$**"

-   $H_{a}:\ p > p_{0}$ (This one is called a **right tail test** )

This one is interpreted "**the population proportion is greater than $p_{0}$**"

-   $H_{a}:\ p \neq p_{0}$ (This one is called a **two tail test** )

This one is interpreted "**the population proportion is not equal to $p_{0}$**"

Which of these you pick depends on what you (as the researcher) want to
show about the population proportion:

-   If you think it is actually smaller than what was claimed, pick the **left tail test**
-   If you think it is actually bigger than what was claimed, pick the **right tail test**
-   If you just want to show its different from what was claimed, and you don't care about how it is different (bigger or smaller would be fine from your viewpoint), pick the **two tailed test**

### 3) Take a (random) sample from the population and compute the sample proportion {-}

$$
\widehat{p} = \frac{count}{\text{sample size}} = \frac{count}{n} 
$$

This is called the **test statistic**

### 4) Compute the z-score for the test statistic {-}

$$
z = \frac{\widehat{p}-p_{0}}{\sqrt{\frac{p_{0}(1-p_{0})}{n}}}
$$

Here you use the (proposed) population proportion $p_{0}$ and the
$\widehat{p}$ is the sample proportion you got in the previous step.

### 5) Find the the P-value {-}

The **P-value** is what you use to conclude your test. 

Its value represents the chance that you get your sample proportion or something more extreme given that the null hypothesis is true.    

In more concrete terms, it is going to be an area (left tail, right tail or both) that comes from your sample proportion.

Here is how you compute it:

-   For **left tail test**, the $P = \text{left tail area}$
-   For **right tail test**, the $P = \text{right tail area}$
-   For **two tail test**, the $P = 2(\text{area of tail})$

To get these areas you just use the z-value from step \#4 above and compute them from that.

### 6a) Find the conclusion from the P-value {-}

To find your conclusion for a hypothesis test once you have your P-value you can use the following guide based on the size of the P-value:

-   P-value bigger than .1, then **little or no** evidence against  ${H_{0}}_{}$
-   P-value between .05 and .1, then **some** evidence against ${H_{0}}_{}$
-   P-value between .01 and .05, then **moderate** evidence against ${H_{0}}_{}$
-   P-value between .001 and .01, then **strong** evidence against ${H_{0}}_{}$
-   P-value less than .001, then **very strong** evidence against ${H_{0}}_{}$

One thing to notice about the conclusions is: **small P-values are
evidence against the null hypothesis.**

That's because a small P-value represents a small area for a tail and
that happens when the z-value (and hence the sample proportion) is more
extreme.

:::{#exm-conclusion-of-hypothesis-test}
## Conclusion of Hypothesis Test
```{asis example=TRUE, title="Conclusion of Hypothesis Test", echo=TRUE}

Suppose you are doing a hypothesis test and you find out that your P-value is $p=.047$.  

Since

$$ 
.01 < .047 < .05 
$$ 

This means that your conclusion is **moderate** evidence against the null $H_0$. 

```
:::

### 6b) Find the conclusion from level of significance ($\alpha$) {-}

Sometimes the conclusion part of a hypothesis test is done in a different way than the above step 6a. 

If you are given a **level of significance** (called $\alpha$) you decide your conclusion in a slightly different way:

- If $P < \alpha$, then you **Reject the null hypothesis**
- If $P \geq \alpha$, then you **Fail to reject the null hypothesis**

Some people use the term **accept the null hypothesis** instead of **fail to reject the null hypothesis**. We take these to be equivalent.  

Using the level of signifcance like this means you state your conclusion differently than in section 6a above. The $\alpha$ is more like a **cut-off** for rejecting the null hypothesis in this case.

And in the case you either **reject the null** or **fail to reject the null**, and there is no othersituation (no "moderate evidence" or "very strong evidence"... just reject or not plain and simple)

:::{#exm-level-of-significance}
## Level of Significance 
```{asis echo=TRUE}

Suppose you are doing a hypothesis test at the 5% level of significance,(which is $\alpha=.05$) and you find out that your P-value is $P=.023$.  

Since

$$ 
.023 < .05 
$$ 

This means that $P < \alpha$ and so you **reject the null hypothesis**. 
```
:::

So in this case you decide your conclusion by comparing to the given $\alpha$ value and not by the method in section 6a. If your P-value is smaller than $\alpha$, you **reject the null hypothesis**, otherwise you **fail to reject the null hypothesis**. 

## Other terminology

Sometimes you will hear people say **accept the null hypothesis** when $p \geq \alpha$. This is just a different way of saying **fail to reject the null hypothesis**.

Finally keep this is mind as well:

- **reject the null hypothesis** means you believe the alternative  
- **failure to reject the null hypothesis** means you believe the null could be true  

And one more that is like a court case verdict:

- **reject the null hypothesis** is in favor of $H_a$  
- **failure to reject the null hypothesis** is in favor of $H_0$  

## Type I and Type II Error

Let us take a minute to mention two types of errors that take place with hypothesis testing called **Type I Error** and **Type II Error.** 

Here are the complete options when you’re dealing with these hypothesis tests. Recall that $H_0$ stands for the null hypothesis. Also we will use the phrase \"accept $H_0$\" instead of \"fail to reject $H_0$\". These mean the same thing. 

- If $H_0$ is actually true and in our testing we accept $H_0$  then there is no mistake
- If $H_0$ is actually false and in our testing we reject $H_0$  then there is no mistake 
- If $H_0$ is actually true and in our testing we reject $H_0$ then that is a mistake and this is called a **Type I Error**
- If $H_0$ is actually false and in our testing we accept $H_0$ then that is a mistake and this is called a **Type II Error**.

Most people give a guilty vs. innocent example so we will too. 

- If an innocent person is not convicted then there is no mistake 
- If a guilty person is convicted then there is no mistake 
- If an innocent person is convicted then that is a mistake and is called a **Type I Error** 
- If a guilty person is not convicted then that is a mistake and is called a **Type II Error**

These errors can impact our daily lives in a huge way, like in the example mentioned above.  More everyday examples include construction projects, our safety, and more. 

## Examples of Hypothesis Testing

The first example is a left tail test:

:::{#exm-left-tail-hypothesis-test-proportions}
## Left Tail Hypothesis Test Proportions
```{r}
knit_example(list(count=190,n=250,p=0.79),
	     'statplosion-problems/question/HypothesisTestPropLeftTailQuestion.Rmd', 
	     'statplosion-problems/solution/HypothesisTestPropLeftTailSolution.Rmd') 
```
:::

Now let's see an example of a right tail test:

:::{#exm-right-tail-hypothesis-test-proportions}
## Right Tail Hypothesis Test-Proportions
```{r}
knit_example(list(count=101,n=250,p=0.34),
	     'statplosion-problems/question/HypothesisTestPropRightTailQuestion.Rmd', 
	     'statplosion-problems/solution/HypothesisTestPropRightTailSolution.Rmd') 
```
:::

## Applications of Hypothesis Testing

Here is an application about selling clothes online:

:::{#exm-selling-clothes-online}
## Selling Clothes Online
```{r}
knit_example(list(count=313,n=700,p=0.40),
	     'book-problems/question/SellingClothesOnlineQuestion.Rmd', 
	     'book-problems/solution/SellingClothesOnlineSolution.Rmd') 
```
:::

Here is one about testing promotion effectiveness: 

:::{#exm-promotion-effectiveness}
## Promotion Effectiveness 
```{r}
knit_example(list(count=38,n=200,p=0.25),
	     'book-problems/question/PromotionEffectivenessQuestion.Rmd', 
	     'book-problems/solution/PromotionEffectivenessSolution.Rmd') 
```
:::
