```{r setup, echo=FALSE}
source("intro-statistics-quarto.R")
library("statplosion")

```

# Straight Line Review {#lines}

In this section we give some examples of straight lines and also look at the interpretation of slope.   

Since we will use straight lines to model data when there is a trend (This is called regression) it is important that we understand what slopes mean and what the equations look like for different lines.  

For slopes:

- a **positive slope** means **_increasing_** or **_uphill trend_**
- a **negative slope** means **_decreasing_** or **_downhill trend_**

Here are some examples:  

**Example 1** 

```{r, fig.width = 4, fig.height=4.4}
m <- 2
b <- 3
plot(0, type="n", xlab="x", ylab="y", xaxs="i", yaxs="i", xlim=c(0, 12), ylim=c(0, 12))
abline(b, m)
```
  
The equation is:  

$y = `r m`x + `r b`$

- The slope is: $m = `r m`$
- intercept is: $b=  `r b`$

**_Interpretation of slope_**

For each additional unit of x, there is an increase of `r m` units of y

**Example 2** 

```{r, fig.width = 4, fig.height=4.4}
m <- -1 
b <- 5 
plot(0, type="n", xlab="x", ylab="y", xaxs="i", yaxs="i", xlim=c(0, 8), ylim=c(0, 8))
abline(5, -1)
```

The equation is:  

$y = `r m`x + `r b`$

- The slope is: $m = `r m`$
- intercept is: $b=  `r b`$

**_Interpretation of slope_**

For each additional unit of x, there is an decrease of `r -m` units of y

**Example 3** 

```{r, fig.width = 4, fig.height=4.4}
m<-2.1
b<-0.5
plot(0, type="n", xlab="advertising expenditures($000s)", ylab="sales($0 000s)", xaxs="i", yaxs="i", xlim=c(0, 8), ylim=c(0, 8))
abline(0.5, 2.1)
```

The equation is:  

$y = `r m`x + `r b`$

- The slope is: $m = `r m`$
- intercept is: $b=  `r b`$

**_Units_**

Suppose $x$ = advertising expenditures (in $000s, so thousands of $)  
Suppose $y$ = sales (in $0 000s, so in ten thousands of $)
    
- So if $x = 1.2$ that stands for $1200   
- If $y = 5.6$ that stands for $56,000     
    
**_Interpretation of slope_**

For each additional unit of x (so each additional $1000 dollars spent on ads), there is an increase of 2.1 units of y (so an increase of $21,000 in sales) 

