```{r setup, echo=FALSE}
source("intro-statistics-quarto.R")
```

# Sampling Distribution of Sample Proportions {#sample-proportions-distribution}

## Repeated Sampling For Proportions

Suppose we start with a population that has a certain population proportion $p$.

Then imagine we take samples of a certain size $n$ over and over again,
each time we calculate the sample proportion $\hat p$.

So we:

- choose a sample and calculate a sample proportion $\hat p$
- choose a sample and calculate another sample proportion $\hat p$
- choose a sample and calculate another sample proportion $\hat p$
- and so on

Usually the sample proportions are close to the population proportion $p$ but sometimes they are farther away (though this is less likely the farther away we get from the population proportion).

**It turns out...**

If we then plot all these sample proportions on a histogram, we get
something that looks like a normal curve!

## Coke vs Pepsi Example

Lets show this with a made up example.

```{r echo=FALSE}
set.seed(2)
n <- 100 
p <- .65
numsamples <- 300 
counts <- rbinom(numsamples,n,p)
phats<-counts/n
```

Suppose the population proportion of students that prefer coke to pepsi is $p=`r p`$. 

This means for the whole student population:

- `r 100*p`\% prefer coke
- `r 100*(1-p)`\% prefer pepsi 

(We don't allow choosing neither as an answer here...its binary)

Suppose we take samples of size $n=`r n`$ students with the following results:

- The first sample has $\hat p = `r phats[1]`$ prefer coke
- The second sample has $\hat p = `r phats[2]`$ prefer coke
- The third sample  has $\hat p = `r phats[3]`$  prefer coke
- and so on

Here are the sample proportions that we get for the first 50 samples:

```{r echo=FALSE, warning=FALSE}
m<-matrix(phats,ncol = 10, byrow=TRUE)
kable(m[1:5,])
```

Suppose we continue and do this `r numsamples` times...

Finally let's plot all of these sample proportions together in a histogram:

```{r echo=FALSE}
hist(phats, breaks=20, 
     xlab="Sample Proportions", 
     ylab="Number", 
     main="Distribution of Sample Proportions That Prefer Coke")
```

We can see that the **shape of the distribution is normal**

## Sampling Distribution of Sample Proportions  

This distribution of the sample proportions is called the **sampling
distribution of sample proportions** or the **$\hat p$-distribution**.

**Sampling distributions** are made by repeatedly taking samples then
computing a statistic, then collecting that data and graphing it in a histogram.

## Central Limit Theorem - Proportions 

It turns out that this "normal curve" shape turns up all the time for
sampling distributions under the right conditions, and that fact is
called the **Central Limit Theorem**.

This is one of the most important results in statistics and basically it
tells us that **we can use a normal distribution to approximate the 
sampling distribution** (the $\hat p$-distribution that is).

This will allow us to calculate areas for the $\widehat{p}$ distribution
by calculating z-values and using a standard normal curve to find areas we want. 
    
:::{#thm-central-limit-theorem}
## Central Limit Theorem-Proportions
```{asis echo=TRUE}
When we use a large population with a population proportion of $p$
and consider samples of size n from this population

- The $\hat p$ distribution is approximately normal, becoming more and more accurate as n is larger  
- The mean and standard deviation of the $\hat p$ distribution are
given by these formulas:

$$
\text{mean} = p
$$

$$
\text{standard deviation} = \sqrt{\frac{p(1-p)}{n}}
$$

- For the above to be valid must have $n(p)>5$ and $n(1-p)>5$.

```
:::
    
The first bullet point of this theorem says:

- we can use a standard normal table to calculate areas under the $\widehat{p}$-distribution

The second bullet point says:

- the $\widehat{p}$-distribution has $p$ as the center of it, and the standard deviation we can compute with the formula in the theorem.

The third bullet point says:

- You have to compute $n(p)$ and $n(1-p)$ and make sure they are at least $5$. Otherwise you cannot use this approach.

## How to Calculate $\hat p$-distribution Areas 

So to find areas under the $\widehat{p}$ distribution we will just
convert them to areas under a standard normal curve like we did before but this time we will use this formula for the z-scores:

$$
z=\frac{\hat p-p}{\sqrt{\frac{(1-p)(p)}{n}}}
$$

So just like before we will find **left tails, right tails** and **areas
between** but this time using the z-value formula above.

## Examples of Areas For $\hat p$-distribution

Lets see some examples of computing areas for the $\hat p$-distribution.

First lets see how to compute a left tail area:

:::{#exm-left-tail-area-for-p-hat-distribution}
## Left Tail Area for $\hat p$-distribution
```{r}
knit_example(list(count=190,n=250,p=0.79),
	     'statplosion-problems/question/SamplingProportionsLeftTailAreaQuestion.Rmd', 
	     'statplosion-problems/solution/SamplingProportionsLeftTailAreaSolution.Rmd') 
```
:::

Now lets see how to compute a right tail area:


:::{#exm-right-tail-area-for-p-hat-distribution}
## Right Tail Area for $\hat p$-distribution
```{r}
knit_example(list(count=208,n=320,p=0.61),
	     'statplosion-problems/question/SamplingProportionsRightTailAreaQuestion.Rmd', 
	     'statplosion-problems/solution/SamplingProportionsRightTailAreaSolution.Rmd') 
```
:::

