Here is the code with the requested find/replace operations:

```{r setup, echo=FALSE}
source("intro-statistics-quarto.R")
```
# Standard Normal Areas {#standard-normal-areas}

In this section we see how to find **left tail areas**, **right tail areas** and also **areas between** for the standard normal distribution by using the z-table. See [Appendix A](#z-table) for the table.

## Standard Normal Left Tail Areas {#standard-normal-left-tail-areas}

The first examples here are for **left tail areas**. We actually already saw how to do this in [Standard Normal Table](#standard-normal-table) by looking up the z-value and reading off the left tail area directly in the table, but it is worth doing one more example since it is so important.

Here is a problem with the z-value is positive:

:::{#exm-Find-left-tail-area-when-z-equals-1.52}

## Find left tail area when $z=1.52$

````{r}
knit_example(list(z=1.52),
             'statplosion-problems/question/StdNormalLeftTailQuestion.Rmd',
             'statplosion-problems/solution/StdNormalLeftTailSolution.Rmd')
````

:::

## Standard Normal Right Tail Areas {#standard-normal-right-tail-areas}

Next lets take a look at finding a **right tail area**. Since our table only uses left tail areas we need a trick. This trick is sometimes called the **1 minus trick**.

It relies on this fact:

$$
\text{right tail area} + \text{left tail area} = 1.0
$$ {#eq-right-tail-plus-left-tail-equals-one}

That is any right tail and its corresponding left tail add up to 100%.

```{r}
z<-1.52
left_area <- pnorm(z)
right_area <- 1.0-left_area
```

So for example below the **right tail is `r round(right_area,2)`** and the **left tail is `r round(left_area,2)`** and these add up to **1.0**

```{r fig.height=2, fig.width=3.5, fig.align="center"}
std_normal_curve()
std_normal_shade_right_area(z)
std_normal_label_point(z)
```

So when we want to find the right tail area for some situation we can just find the left tail area and then subtract. Here are some examples.

:::{#exm-Find-right-tail-area-when-z-equals-minus-1.32}

## Find right tail area when $z=-1.32$

````{r}
knit_example(list(z=-1.32),
             'statplosion-problems/question/StdNormalRightTailQuestion.Rmd',
             'statplosion-problems/solution/StdNormalRightTailSolution.Rmd')
````

:::

Another example:

:::{#exm-Find-right-tail-area-when-z-equals-0.52}

## Find right tail area when $z=0.52$

````{r}
knit_example(list(z=0.52),
             'statplosion-problems/question/StdNormalRightTailQuestion.Rmd',
             'statplosion-problems/solution/StdNormalRightTailSolution.Rmd')
````

:::


## Standard Normal Area Between Two Z Values {#standard-normal-area-between}

Finding the **area between** involves a trick as well.

_**Find Area Between Z Values**_

1. You look up the **left tail area** to the "upper z" (the one farthest to the right).
2. Then you look up the **left tail area** to the "lower z" (the one farthest to the left).
3. Then you subtract those two areas to get the **area between**.

Be sure to subtract these the right way. If you subtract them backwards your answer will turn out negative. Since areas cannot be negative, you will know you made a mistake.

Okay here are some examples of finding the **area between** two z-values.

:::{#exm-Find-area-between-z_1-equals-0.52-and-z_2-equals-1.65}

## Find area between $z_1=0.52$ and $z_2=1.65$

````{r}
knit_example(list(z1=0.52, z2=1.65),
             'statplosion-problems/question/StdNormalAreaBetweenQuestion.Rmd',
             'statplosion-problems/solution/StdNormalAreaBetweenSolution.Rmd')
````

:::

Another one

:::{#exm-Find-area-between-z_1-equals-minus-1.53-and-z_2-equals-1.23}

## Find area between $z_1=-1.53$ and $z_2=1.23$

````{r}
knit_example(list(z1=-1.53, z2=1.23),
             'statplosion-problems/question/StdNormalAreaBetweenQuestion.Rmd',
             'statplosion-problems/solution/StdNormalAreaBetweenSolution.Rmd')
````

:::

## Applications of Standard Normal Distributions

We will see later in the text that **_service level_** in retail corresponds to the chance that a retailer would experience a lost sale (out of stock situation) for some period of interest.

- A service level of 90% means that lost sales would happen 10% of the time.
- A service level of 95% means that lost sales would happen be 5% of the time.
- A service level of 99% means that lost sales would happen be 1% of the time.

For demand that is distributed normally, the service level can be represented as a **left tail area**.

In addtion, the $z$-value that goes with a particular service level is called the **_service factor_**.

|   |   |
|---|---|
| service level | left tail area |
| service factor  | $z$-value for that left tail area |

- If you know the service factor ($z$-value) you can find the service level (the left tail area). (this section)
- If you know the service level (left tail area), you can find the service factor ($z$-value) (later in book)

:::{#exm-Service-Level}

## Service Level

````{r}
knit_example(list(z=1.28),
             'book-problems/question/ServiceLevelQuestion.Rmd',
             'book-problems/solution/ServiceLevelSolution.Rmd')
````

:::
