```{r setup, echo=FALSE}
source("intro-statistics-quarto.R")
```
# Standard Normal Table {#standard-normal-table}

This section shows how to use the standard normal table to find left tail areas. We will be finding the areas for examples like [these](#left-tail-areas).  

The standard normal table (z-table) we use is in [Appendix A](#z-table) This table contains *only* **left tail areas** in it, so in this section we concentrate on showing examples of finding **left tail area** using the table. 

If you have a **right tail area** or **area between** that you need to find, you will have to use some tricks to find the area you want by using the table and some left tail area. In [Standard Normal Areas](#standard-normal-areas) we see how to do that.

So here are a few examples of finding **left tail areas** using the table:

:::{#exm-Left-tail-area-for-$z=1.25$}
## Left tail area for $z=-1.25$ 
```{r}
knit_example(list(z=-1.25), 'TableLookupQuestion.Rmd', 'TableLookupSolution.Rmd')
```
:::

:::{#exm-Left-tail-area-for-$z=1.32$}
## Left tail area for $z=1.32$ 
```{r}
knit_example(list(z=1.32), 'TableLookupQuestion.Rmd', 'TableLookupSolution.Rmd')
```
:::

:::{#exm-Left-tail-area-for-$z=0.07$}
## Left tail area for $z=0.07$ 
```{r}
knit_example(list(z=0.07), 'TableLookupQuestion.Rmd', 'TableLookupSolution.Rmd')
```
:::
