```{r setup, echo=FALSE}
source("intro-statistics-quarto.R")
```
# My Chapter {#sec-my-chapter}

Let's take at a look at an important calculation: Finding $z$-values

| fruit  | price  |
|--------|--------|
| apple  | 2.05   |
| pear   | 1.37   |
| orange | 3.09   |

: Fruit prices {#tbl-fruit-prices .striped .hover}

You can do the link to the above this: @tbl-fruit-prices. 
Or you can do it like this: [Fruit Prices](test.Rmd#tbl-fruit-prices)


| Col1 | Col2 | Col3 |
|------|------|------|
| A    | B    | C    |
| E    | F    | G    |
| A    | G    | G    |

: My Caption {.striped}

If you do not give it a label you cannot reference it with a link. And it does not get the table numbering.

::: {style="width:33%;margin:auto"}
```{r tbl-cars}
#| tbl-cap: These are the cars! 

kable(head(cars))
```
:::

You can use this calculation to figure out how far away data values are from the mean:


Here is another example:

::: {style="width:33%;margin:auto"}
```{asis tbl-fruits-and-prices, echo=T}
#| tbl-cap: These are the fruits and prices 

| fruit  | price  |
|--------|--------|
| apple  | 2.05   |
| pear   | 1.37   |
| orange | 3.09   |
```
:::

| |what| symbol  | when to use  |
|---|---|---|----------|
| population mean | parameter | $\mu$  | for mean of entire population  |
| sample mean  | statistic | $\bar x$   | for mean of sample |

Here is the formula in words:

$$
z =\frac{\text{data} -\text{mean}}{\text{std dev}}
$$ {#eq-z-value-formula-in-words}


## Population Z Value

First lets look at the population version of it. Suppose you know the mean $\mu$ and the standard deviation $\sigma$ for some population. You can find the $z$-value for any data value $x$ like this:

$$
z =\frac{x -\mu}{\sigma}
$$ {#eq-z-value-formula}

Lets see some examples of this in action:
adfd

:::{#exm-Find-the-z-value}
## Find the z-value
```{r, fig.label="wee", fig.cap="weee"}
knit_example(list(mu=52, x=54, sigma=8),
	     "statplosion-problems/question/ZValueQuestion.Rmd",
	     "statplosion-problems/solution/ZValueSolution.Rmd")
```
:::

asdfaf

Here is an example where we find the z-value of an exam score of $65$:

:::{#exm-Left-tail-area-for-z-equals-minus-1.02-using-spreadsheet}
## Left tail area for $z=-1.02$ using spreadsheet
```{r}
knit_example(list(z=-1.02),
             'statplosion-problems/question/StdNormalLeftTailSpreadsheetQuestion.Rmd',
             'statplosion-problems/solution/StdNormalLeftTailSpreadsheetSolution.Rmd')
```
:::
